Download data

### 2024-05-03 scripts/GroundTruthSampling.Rmd
dispatch

```{r import study extent and convert polygons to rasters}

@@

```{r mask areas with fewer predicted species}
samples <- spatSample(r, 10000, na.rm = T)

lyr1 <- as.numeric(quantile(samples$SumOccupiedPatches, probs = 0.75))
lyr2 <- as.numeric(quantile(samples$Sum1stOrderNeighboringPatches, probs = 0.75))
lyr3 <- as.numeric(quantile(samples$CommonessInversion, probs = 0.75))

@@

```{r}
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)
```

@@

lapply(ob, identify_unqiue_pts)
```

>>>>

```{r import study extent and convert polygons to rasters, eval = F}

@@

```{r mask areas with fewer predicted species}

r <- rast( '../results/Summary_Rasters/COPL.tif')
samples <- spatSample(r, 10000, na.rm = T)

lyr1 <- as.numeric(quantile(samples$SumOccupiedPatches, probs = 0.5))
lyr2 <- as.numeric(quantile(samples$Sum1stOrderNeighboringPatches, probs = 0.5))
lyr3 <- as.numeric(quantile(samples$CommonessInversion, probs = 0.5))

@@

```{r reload data to consider each species independently}
datr <- lapply(dat, rasterize, r, field = 'Rank')
datr <- rast(datr)
```

We will consider three attributes of each sampled population...

```{r draw random points for each species}

@@

ob1 <- lapply(ob, identify_unqiue_pts) |>
  bind_rows()
```

Let's determine which sites have useful info

obbies <- ob1 %>% 
  rowwise() %>% 
  mutate(sum_na = sum(is.na(c_across(SumOccupiedPatches:CommonessInversion)))) %>% 
  group_by(taxon, Target) %>% 
  arrange(taxon, Target, sum_na, -SumOccupiedPatches, -Sum1stOrderNeighboringPatches, -CommonessInversion) %>% 
  slice_head(n = 3)

```

Create site indexes for all sample points

ggplot() + 
  geom_sf(data = obbies)

disties <- st_nearest_feature(obbies)  
dd <- st_distance(obbies, obbies[disties,], by_element = TRUE)

p <- '/media/steppe/hdd/2024SOS_CrewGeospatial/species_check_play/'
st_write(obbies, dsn = paste0(p, 'test.shp'))

```

### 2024-05-03 scripts/subsetDataforCrews.Rmd
dispatch

```{r jitter the AIM data, eval = F}
set.seed(125)
occ_p <- '/media/steppe/hdd/SDM_restorations/data/raw/occurrence/combined_records'
occurrences <- lapply(
  file.path(occ_p, list.files(occ_p, pattern = 'shp$')), st_read, quiet = T) |>
  bind_rows() %>% 
  filter(Occurrence == 1) %>% 
  select(taxon, date, PtType)
occ_aim <- filter(occurrences, PtType == 'AIM Presence') %>% 
  st_jitter(amount = 250)
occurrences <- bind_rows(
  filter(occurrences, PtType != 'AIM Presence'), 
  occ_aim) %>% 
  mutate(
    PtType = if_else(PtType == 'AIM Presence', 'jittered', 'exact'),
    date = as.numeric(str_extract(date, '[0-9]{4}'))
    ) %>% 
  rename(year = date, coord_status = PtType) %>% 
  arrange(taxon, year, coord_status)

@@

```{r seed zones, eval = F}
seed_zones <- read.csv('../data/raw/2024sos_species-FIXED.csv') %>% 
    unite('binomial', genus:species)

seed_zones %>% 
  filter(stz == 'BowerRevised') %>% 
  distinct(ind_stz) %>% 
  mutate(
    ind_stz = str_replace(ind_stz, '  ', ' ')
  ) %>% 
  separate_wider_delim(ind_stz, delim = ',', names= LETTERS[1:7], too_few = "align_start")

  #  ind_stz = str_replace(ind_stz, ' -[0-9]|[0-9]- |[0-9]-[0-9]', ' - ') )
```

@@

@@

```{r}
teams_list <- split(teams, f = teams$Contract)
```

@@

lapply(teams_list, project_maker,
       admu = admin_boundaries, 
       target_species = spp_needed, 
       blm_surf = blm_surf, 
       fs_surf = usfs_surf, 
       fire = fire_perimeters, 
       invasive = invasives, 
       total_change = change_intensity, 
       occurrences = occurrences, 
      # historic_SOS, 
       roads = roads, seed_transfer = STZ)

table(STZ$Dataset) 
head(STZ)
```

@@

p2ts <- '/media/steppe/hdd/SeedPhenology/data/processed/timestamps'
ott1 <- lapply(
  X = teams_list, # 5 throws an error
  FUN = phen_tabulator,
  path = p2ts, 
  admu = admin_boundaries, 
  target_species = spp_needed
  ) # row 67

>>>>

```{r jitter the AIM data, eval = F}

set.seed(125)
occ_p <- '/media/steppe/hdd/SDM_restorations/data/raw/occurrence/combined_records'
occurrences <- lapply(
  file.path(occ_p, list.files(occ_p, pattern = 'shp$')), st_read, quiet = T) |>
  bind_rows() %>% 
  filter(Occurrence == 1) %>% 
  select(taxon, date, PtType)
occ_aim <- filter(occurrences, PtType == 'AIM Presence') %>% 
  st_jitter(amount = 250)
occurrences <- bind_rows(
  filter(occurrences, PtType != 'AIM Presence'), 
  occ_aim) %>% 
  mutate(
    PtType = if_else(PtType == 'AIM Presence', 'jittered', 'exact'),
    date = as.numeric(str_extract(date, '[0-9]{4}'))
    ) %>% 
  rename(year = date, coord_status = PtType) %>% 
  arrange(taxon, year, coord_status)

@@

```{r seed zones}
seed_zones <- read.csv('../data/raw/2024sos_species-FIXED.csv') %>% 
    unite('binomial', genus:species)

seed_zones <- seed_zones %>% 
  mutate(
    ind_stz = str_replace(ind_stz, '  ', ' ')
  ) %>% 
  mutate(commas = max(str_count(ind_stz, pattern = ','))+1) %>% 
  separate_wider_delim(ind_stz, delim = ',', names= LETTERS[1:.$commas], too_few = "align_start") %>% 
  mutate(across(everything(), ~ str_trim(.))) %>% 
  relocate(contract, .before = 'A') %>% 
  pivot_longer(cols = LETTERS[1:.$commas], values_to = 'st_zone') %>% 
  select(-commas, -name) %>% 
  drop_na(st_zone) %>% 
  mutate(stz = case_when(
    stz == 'Desert SW provisional' ~ 'DesertSW', 
    stz == 'BowerRevised' ~ 'Bower2013', 
    .default = as.character(stz)
  ))

seed_zones <- read.csv('../data/raw/contract_lkp.csv') %>% 
  left_join(seed_zones, ., by = c('contract' = 'Contract_colloquial'))

@@

prov_stz <- rast('/media/steppe/hdd/EmpiricalSeedZones/data/geodata/final_provisional/provisional_stack.tif')

@@

```{r}
teams_list <- split(teams, f = teams$Contract)
seniors_list <- split(seniors, f = seniors$Senior)

t <- filter(occurrences, taxon == 'Abronia villosa') |>
  vect() |>
  project(crs(prov_stz))

class(t)[1] == 'sf' 
unique(st_geometry_type(t))[1] == 'POINT'

# need to be able to dispatch to both raster and vector data, and both point and
# polygon data. 
if(class(x) == 'SpatRaster'){

  # dispatch method for raster

} else if(class(x) == 'sf'){ # determine type of vector data and dispatch.

  if(unique(sf::st_geometry_type(x))[1] == 'POINT'){

  } else {

  }
} else {
  message('This class not expected by this small function. Small overhaul needed?')
  }

# this identified the occurrence records which we want to keep. 
which(
  !is.na(
    terra::extract(prov_stz[[2]], t, method = 'simple', ID = F)))

@@

unique(seed_zones$stz)

lapply(teams_list[13], project_maker,
       admu = admin_boundaries, 
       target_species = spp_needed, 
       blm_surf = blm_surf, 
       fs_surf = usfs_surf, 
       fire = fire_perimeters, 
       invasive = invasives, 
       total_change = change_intensity, 
       occurrences = occurrences, 
      # historic_SOS, 
       roads = roads, seed_transfer = STZ)

lapply(seniors_list[3],  project_maker,
       admu = admin_boundaries, intype = 'senior',
       target_species = spp_needed, 
       blm_surf = blm_surf, 
       fs_surf = usfs_surf, 
       fire = fire_perimeters, 
       invasive = invasives, 
       total_change = change_intensity, 
       occurrences = occurrences, 
      # historic_SOS, 
       roads = roads, seed_transfer = STZ)
```

@@

p2ts <- '/media/steppe/hdd/SeedPhenology/data/processed/timestamps'
lapply(
  X = teams_list[13], # 5 throws an error
  FUN = phen_tabulator,
  path = p2ts, 
  admu = admin_boundaries, 
  target_species = spp_needed
  ) # row 67
